SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-11 docs/questioning.tex
SB becomes StB and EQ becomes EiQ

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source
code injection vulnerabilities. For every arrow there is a
corresponding tuple that includes a transition number for
every attack. Every {\it point} indicates a transition where mechanisms
detect attacks. For every point we provide the corresponding
references:
Point {\sc ub}~\cite{KJKV09,TNH07,NSS06,APKLM10,SPPGMMB04,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,GC09,VFJKKV07,SLMS14,BV08},
Point {\sc sb}~\cite{RDWDE07,JKK06a,JB07,NLC07,WPLKK09,JEP08},
Point {\sc eq}~\cite{BWS05,SW06,PS11,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point {\sc sdb}~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

Detection mechanisms vary on the location where they detect an attack.
There are four different points where an attack can be caught, as seen
in Figure~\ref{fig:attacks}:
\begin{itemize}
\item At the user's browser (Point {\sc ub}).
\item En route from the server to the user's browser --- in most cases
within a proxy (Point {\sc sb}).
\item Embedded in a query, but before it reaches the server's database
  (Point {\sc eq}).
\item After the malicious code has been executed on the server's
  database (Point {\sc sdb}).
\end{itemize}

@@

In the case of {\sc dsl} injection attacks, mechanisms check 
the query before the inclusion of user input with the one
resulting after the inclusion of user input.
Both mechanisms that implement this approach,
{\sc sqlg}uard~\cite{BWS05} and
{\sc sql}check~\cite{SW06} are pretty similar
and detect the attack before a query reaches the
database (Point {\sc eq}).
Contrary to {\sc sqlg}uard (and other
mechanisms as we will observe), {\sc sql}check has been
extensively tested in terms of accuracy
as it can be seen in Table~\ref{tab:comp}.
A disadvantage of those mechanisms is that the
application must be modified in every code fragment
that sends an {\sc sql} query to the database for execution.

@@

Another policy enforcement approach introduces
policies directly either in {\sc html} or JavaScript code
to confine their behavior. {\it BrowserShield}~\cite{RDWDE07}
acts as a proxy on the server side (Point {\sc SB}) to
parse the {\sc html} of server responses and identifies
scripts. Then, it rewrites them into safe equivalents
and protects the web user from exploits
that are based on reported browser vulnerabilities.
{\it ConScript}~\cite{ML10}, {\it CoreScript}~\cite{YCIS07}
and the framework by {\it Phung et al.}~\cite{PSC09}
extend JavaScript with new primitive functions that
provide safe methods to protect potentially vulnerable
JavaScript functions. In both cases, policy enforcement takes
place on the JavaScript engine of the browser (Point {\sc ub}).
In this way, {\sc xxs} attacks that take advantage
functions like {\tt write} and {\tt eval}, which are
used to assemble innocuous-looking parts into harmful
strings,\footnote{The infamous Sammy worm that
infected MySpace in 2005~\cite{SP07,ELX07}
utilized the {\tt eval} function to assemble a
script with malicious behavior.} would fail.

@@

Policy enforcement mechanisms that detect {\sc csrf}
attacks are usually implemented in the form of a
server-side proxy (Point {\sc SB})
interposing on the client-server communication, and modifying
it. {\it NoForge}~\cite{JKK06a}
parses the {\sc html} server responses
and adds a token to every {\sc url} referring to this
server. Then it associates the token with the cookie
representing the session {\sc id} for the application.
When a request is received, the mechanism checks
if the request contains the token related
to the session {\sc id}. A disadvantage of NoForge
is that {\sc html} that is dynamically created within
the browser will not include the token.
Thus, sites that create their {\sc html} on the client
will remain vulnerable. In addition, it does not
support cross-origin requests.
{\it j{\sc csrf}}~\cite{PS11}
has a very similar functionality and meanwhile
addresses the above problems.
Finally, {\it CsFire}~\cite{DDHPJ10}
examines cross-domain interactions, to design a
cross-domain policy on the client's side (Point {\sc ub}).
The policy is based on the concept of a relaxed
same-origin policy that allows communication between
sub-domains of the same registered domain.

@@

In the case of {\sc xss} attacks that are based on JavaScript or {\sc
  html}, consider a {\sc xor} function that encodes all source of a
web page on the server-side and then, on the client-side, the web
browser decodes the source by applying the same function again (Point
{\sc ub}). Variations of this approach include {\it
  Noncespaces}~\cite{GC09} and {\it x{\sc js}}~\cite{APKLM10}, which
randomize the instruction set of {\sc html} and JavaScript
respectively. Contrary to x{\sc js}, in Noncespaces administrators
must set specific policies in a manner similar to a firewall
configuration language. This fact, together with the insufficient
testing of the mechanism (see Table~\ref{tab:comp}) brings up
questions concerning its accuracy. {\it {\sc
    sm}ask}~\cite{JB07} is another framework that was inspired by {\sc
  isr}. To detect {\sc xss} attacks it searches for {\sc html} and
JavaScript keywords within the application's legitimate code. This is
done before the processing of an {\sc http} request. When it finds one
it adds a token to it. This results to a ``code mask". Then, before
sending the resulting {\sc html} data to the web user, the framework
searches the data for illegal code by using the same keywords (Point
{\sc SB}). Since all legitimate code has been ``masked", the injected
code can be identified. The pre- and post-processing of the code
though, may add a significant overhead to the application.
Unfortunately, the authors of {\sc sm}ask did not provide measurements
regarding the computational performance of the tool (see
Table~\ref{tab:comp}).

@@

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (Point {\sc eq}).

It is possible to apply further checks when it is established that
tainted data have reached a sink (Point {\sc eq}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Point {\sc eq}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of {\sc php} arrays data structure.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (Point {\sc sb}).

@@

Training methods that detect {\sc dsl}-driven injection attacks
generate and store valid code statements (i.e. {\sc sql} or {\sc
  xp}ath queries) in various forms, and detect attacks as outliers
from the set of valid code statements. An early approach, {\it {\sc
    didafit}}~\cite{LLW02} detects {\sc sql} injection (Point
sdb) attacks by recording all database transactions
(stripped from user input). Subsequent
refinements by Valeur et al.~\cite{VMV05} tagged each transaction with
the corresponding application as an extension of their anomaly
detection framework called {\it libAnomaly}. {\it {\sc
    sd}river}~\cite{MS09,MKS09,MKLS11} is a signature-based mechanism
that prevents {\sc sql} and {\sc xp}ath injection attacks. The
signatures generated during a training phase are based on features
that can depend either on the code statement or on its execution
environment (i.e. the stack trace). Then, at runtime, the mechanism
checks all statements for compliance and can block code statements
containing injected elements (Point {\sc eq}). {\sc
  amnesia}~\cite{HO05,HO06,HO05b} is a tool that also detects {\sc
  sql} injection attacks (Point {\sc eq}) by associating a query model
with the location of every {\sc sql} statement within the application.
Then, at runtime, it monitors the application's execution to detect
when {\sc sql} statements diverge from the expected model.

Various countermeasures against {\sc xss} attacks follow a similar
pattern. {\it {\sc swap}}~\cite{WPLKK09} creates a unique identifier
({\it script {\sc id}}) for every legitimate script on the server.
Then, a JavaScript detection component placed in a web proxy (Point
{\sc sb}) searches for injected scripts with no corresponding {\sc id}
in the server's responses. If no injected scripts are found, the proxy
forwards the request to the client. This mechanism is relatively
inflexible since it does not support dynamic scripts. In addition it
imposes a significant overhead (see Table~\ref{tab:comp}). The authors
of {\it {\sc xssds}}~\cite{JEP08} have implemented a similar mechanism
that also supports dynamic and external scripts. Specifically, during
training phase they build a list of all benign scripts. In the case of
external scripts they keep a white list of all the valid domain names
that contain scripts used by the application.

@@

{\it Diglossia}~\cite{SMS13} combines positive taint tracking (see
Subsubsection~\ref{sec:taint}) together with parse-tree validation.
Diglossia was based on the theory of Ray and Ligatti~\cite{RL12b}
(also see Section~\ref{sec:attacks}) to detect {\sc dsl} code
injection attacks. In addition, it is actually the first, and so far
the only framework that detects No{\sc sql} injection attacks. When an
application computes an output string (query), Diglossia computes a
``shadow" of this string. Specifically, it maps all characters
introduced by the application to a shadow character set. This set does
not contain any characters coming from the tainted input. Then, the
scheme creates the tree of the query that is about to be executed and
compares it with the parsed tree of the ``shadow" (Point {\sc eq}). If
the trees are not equal the application is probably under attack. Note
that Diglossia can be bypassed in the same way as other taint
tracking approaches as described in reference~\cite{NBR14}.
%% PL which way is this exactly?

@@

Given the fact that the final steps of
{\sc dsl} injection and {\sc csrf} attacks take place
on the server-side, all the mechanisms that detect
such attacks are placed in some point within the server.
From the mechanisms that detect {\sc dsl} code injection
attacks, 3 out of 13 (23\%) do so at the level
of the {\sc rdbms}---Point {\sc sdb}.
The other 11 wrap {\sc api} calls related to output
vectors---Point {\sc eq}.

>>>>

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source
code injection vulnerabilities. For every arrow there is a
corresponding tuple that includes a transition number for
every attack. Every {\it point} indicates a transition where mechanisms
detect attacks. For every point we provide the corresponding
references:
Point {\sc ub}~\cite{KJKV09,TNH07,NSS06,APKLM10,SPPGMMB04,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,GC09,VFJKKV07,SLMS14,BV08},
Point {\sc s}t{\sc b}~\cite{RDWDE07,JKK06a,JB07,NLC07,WPLKK09,JEP08},
Point {\sc e}i{\sc q}~\cite{BWS05,SW06,PS11,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point {\sc sdb}~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

Detection mechanisms vary on the location where they detect an attack.
There are four different points where an attack can be caught, as seen
in Figure~\ref{fig:attacks}:
\begin{itemize}
\item At the user's browser (Point {\sc ub}).
\item En route from the server to the user's browser --- in most cases
within a proxy (Point {\sc s}t{\sc b}).
\item Embedded in a query, but before it reaches the server's database
  (Point {\sc e}i{\sc q}).
\item After the malicious code has been executed on the server's
  database (Point {\sc sdb}).
\end{itemize}

@@

In the case of {\sc dsl} injection attacks, mechanisms check 
the query before the inclusion of user input with the one
resulting after the inclusion of user input.
Both mechanisms that implement this approach,
{\sc sqlg}uard~\cite{BWS05} and
{\sc sql}check~\cite{SW06} are pretty similar
and detect the attack before a query reaches the
database (Point {\sc e}i{\sc q}).
Contrary to {\sc sqlg}uard (and other
mechanisms as we will observe), {\sc sql}check has been
extensively tested in terms of accuracy
as it can be seen in Table~\ref{tab:comp}.
A disadvantage of those mechanisms is that the
application must be modified in every code fragment
that sends an {\sc sql} query to the database for execution.

@@

Another policy enforcement approach introduces
policies directly either in {\sc html} or JavaScript code
to confine their behavior. {\it BrowserShield}~\cite{RDWDE07}
acts as a proxy on the server side (Point {\sc s}t{\sc b}) to
parse the {\sc html} of server responses and identifies
scripts. Then, it rewrites them into safe equivalents
and protects the web user from exploits
that are based on reported browser vulnerabilities.
{\it ConScript}~\cite{ML10}, {\it CoreScript}~\cite{YCIS07}
and the framework by {\it Phung et al.}~\cite{PSC09}
extend JavaScript with new primitive functions that
provide safe methods to protect potentially vulnerable
JavaScript functions. In both cases, policy enforcement takes
place on the JavaScript engine of the browser (Point {\sc ub}).
In this way, {\sc xxs} attacks that take advantage
functions like {\tt write} and {\tt eval}, which are
used to assemble innocuous-looking parts into harmful
strings,\footnote{The infamous Sammy worm that
infected MySpace in 2005~\cite{SP07,ELX07}
utilized the {\tt eval} function to assemble a
script with malicious behavior.} would fail.

@@

Policy enforcement mechanisms that detect {\sc csrf}
attacks are usually implemented in the form of a
server-side proxy (Point {\sc s}t{\sc b})
interposing on the client-server communication, and modifying
it. {\it NoForge}~\cite{JKK06a}
parses the {\sc html} server responses
and adds a token to every {\sc url} referring to this
server. Then it associates the token with the cookie
representing the session {\sc id} for the application.
When a request is received, the mechanism checks
if the request contains the token related
to the session {\sc id}. A disadvantage of NoForge
is that {\sc html} that is dynamically created within
the browser will not include the token.
Thus, sites that create their {\sc html} on the client
will remain vulnerable. In addition, it does not
support cross-origin requests.
{\it j{\sc csrf}}~\cite{PS11}
has a very similar functionality and meanwhile
addresses the above problems.
Finally, {\it CsFire}~\cite{DDHPJ10}
examines cross-domain interactions, to design a
cross-domain policy on the client's side (Point {\sc ub}).
The policy is based on the concept of a relaxed
same-origin policy that allows communication between
sub-domains of the same registered domain.

@@

In the case of {\sc xss} attacks that are based on JavaScript or {\sc
  html}, consider a {\sc xor} function that encodes all source of a
web page on the server-side and then, on the client-side, the web
browser decodes the source by applying the same function again (Point
{\sc ub}). Variations of this approach include {\it
  Noncespaces}~\cite{GC09} and {\it x{\sc js}}~\cite{APKLM10}, which
randomize the instruction set of {\sc html} and JavaScript
respectively. Contrary to x{\sc js}, in Noncespaces administrators
must set specific policies in a manner similar to a firewall
configuration language. This fact, together with the insufficient
testing of the mechanism (see Table~\ref{tab:comp}) brings up
questions concerning its accuracy. {\it {\sc
    sm}ask}~\cite{JB07} is another framework that was inspired by {\sc
  isr}. To detect {\sc xss} attacks it searches for {\sc html} and
JavaScript keywords within the application's legitimate code. This is
done before the processing of an {\sc http} request. When it finds one
it adds a token to it. This results to a ``code mask". Then, before
sending the resulting {\sc html} data to the web user, the framework
searches the data for illegal code by using the same keywords (Point
{\sc s}t{\sc b}). Since all legitimate code has been ``masked", the injected
code can be identified. The pre- and post-processing of the code
though, may add a significant overhead to the application.
Unfortunately, the authors of {\sc sm}ask did not provide measurements
regarding the computational performance of the tool (see
Table~\ref{tab:comp}).

@@

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (Point {\sc e}i{\sc q}).

It is possible to apply further checks when it is established that
tainted data have reached a sink (Point {\sc e}i{\sc q}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Point {\sc e}i{\sc q}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of {\sc php} arrays data structure.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (Point {\sc s}t{\sc b}).

@@

Training methods that detect {\sc dsl}-driven injection attacks
generate and store valid code statements (i.e. {\sc sql} or {\sc
  xp}ath queries) in various forms, and detect attacks as outliers
from the set of valid code statements. An early approach, {\it {\sc
    didafit}}~\cite{LLW02} detects {\sc sql} injection (Point
sdb) attacks by recording all database transactions
(stripped from user input). Subsequent
refinements by Valeur et al.~\cite{VMV05} tagged each transaction with
the corresponding application as an extension of their anomaly
detection framework called {\it libAnomaly}. {\it {\sc
    sd}river}~\cite{MS09,MKS09,MKLS11} is a signature-based mechanism
that prevents {\sc sql} and {\sc xp}ath injection attacks. The
signatures generated during a training phase are based on features
that can depend either on the code statement or on its execution
environment (i.e. the stack trace). Then, at runtime, the mechanism
checks all statements for compliance and can block code statements
containing injected elements (Point {\sc e}i{\sc q}). {\sc
  amnesia}~\cite{HO05,HO06,HO05b} is a tool that also detects {\sc
  sql} injection attacks (Point {\sc e}i{\sc q}) by associating a query model
with the location of every {\sc sql} statement within the application.
Then, at runtime, it monitors the application's execution to detect
when {\sc sql} statements diverge from the expected model.

Various countermeasures against {\sc xss} attacks follow a similar
pattern. {\it {\sc swap}}~\cite{WPLKK09} creates a unique identifier
({\it script {\sc id}}) for every legitimate script on the server.
Then, a JavaScript detection component placed in a web proxy (Point
{\sc s}t{\sc b}) searches for injected scripts with no corresponding {\sc id}
in the server's responses. If no injected scripts are found, the proxy
forwards the request to the client. This mechanism is relatively
inflexible since it does not support dynamic scripts. In addition it
imposes a significant overhead (see Table~\ref{tab:comp}). The authors
of {\it {\sc xssds}}~\cite{JEP08} have implemented a similar mechanism
that also supports dynamic and external scripts. Specifically, during
training phase they build a list of all benign scripts. In the case of
external scripts they keep a white list of all the valid domain names
that contain scripts used by the application.

@@

{\it Diglossia}~\cite{SMS13} combines positive taint tracking (see
Subsubsection~\ref{sec:taint}) together with parse-tree validation.
Diglossia was based on the theory of Ray and Ligatti~\cite{RL12b}
(also see Section~\ref{sec:attacks}) to detect {\sc dsl} code
injection attacks. In addition, it is actually the first, and so far
the only framework that detects No{\sc sql} injection attacks. When an
application computes an output string (query), Diglossia computes a
``shadow" of this string. Specifically, it maps all characters
introduced by the application to a shadow character set. This set does
not contain any characters coming from the tainted input. Then, the
scheme creates the tree of the query that is about to be executed and
compares it with the parsed tree of the ``shadow" (Point {\sc e}i{\sc q}). If
the trees are not equal the application is probably under attack. Note
that Diglossia can be bypassed in the same way as other taint
tracking approaches as described in reference~\cite{NBR14}.
%% PL which way is this exactly?

@@

Given the fact that the final steps of
{\sc dsl} injection and {\sc csrf} attacks take place
on the server-side, all the mechanisms that detect
such attacks are placed in some point within the server.
From the mechanisms that detect {\sc dsl} code injection
attacks, 3 out of 13 (23\%) do so at the level
of the {\sc rdbms}---Point {\sc sdb}.
The other 11 wrap {\sc api} calls related to output
vectors---Point {\sc e}i{\sc q}.
